---
title: "all_strain_genome_stats"
output: pdf_document
date: "2023-07-12"
---
## Assembly stats for 15 Emiliania huxleyi genomes

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(stringr)
library(ggrepel)
library(plyr)
library(cowplot)
```

#### Read in stats for pre-decontamination assemblies
```{r read_in_contam}
genstats <- list()
gc <- list()
contam_stats <- list()
strains <- as.character(read.csv('data/strains.csv',header=FALSE))
for (strain in strains){
  genstats[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_no_pilon_contaminated/"
                                          ,strain,".genstats.txt"),sep="\t",header=F)
  colnames(genstats[[strain]]) <- c(	'Contig_name','Avg_fold',	'Length',	'Ref_GC',	
                                     'Covered_percent',	'Covered_bases','Plus_reads',
                                     'Minus_reads',	'Read_GC',	'Median_fold', 'Std_Dev')

  gc[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_no_pilon_contaminated/"
                                      ,strain,".gcscaffold.txt"),sep="\t",header=F)
  colnames(gc[[strain]]) <- c('Contig_name',	'Length',	'A',	'C',	'G',	'T',	'N',	
                              'IUPAC','Other',	'GC')

    #genstats[[strain]]$Contig_name <- str_split(genstats[[strain]]$Contig_name,regex("_1$"),simplify = T)[,1]
    
  contam_stats[[strain]] <- inner_join(genstats[[strain]],gc[[strain]],by="Contig_name")


}
```


#### Read in assembly stats data for non-dedupped assemblies

```{r read_in}
genstats <- list()
gc <- list()
dup_stats <- list()
for (strain in strains){
  if (file.exists(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam/"
                                          ,strain,".genstats.txt"))){
  genstats[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam/"
                                          ,strain,".genstats.txt"),sep="\t",header=F)
  colnames(genstats[[strain]]) <- c(	'Contig_name','Avg_fold',	'Length',	'Ref_GC',	
                                     'Covered_percent',	'Covered_bases','Plus_reads',
                                     'Minus_reads',	'Read_GC',	'Median_fold', 'Std_Dev')

  gc[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam/"
                                      ,strain,".gcscaffold.txt"),sep="\t",header=F)
  colnames(gc[[strain]]) <- c('Contig_name',	'Length',	'A',	'C',	'G',	'T',	'N',	
                              'IUPAC','Other',	'GC')

    #genstats[[strain]]$Contig_name <- str_split(genstats[[strain]]$Contig_name,regex("_1$"),simplify = T)[,1]
    
  dup_stats[[strain]] <- inner_join(genstats[[strain]],gc[[strain]],by="Contig_name")
}
else{
  print(strain)
}
}
```
#### Read in stats for dedup assemblies

```{r read_in_rmdup}
genstats <- list()
gc <- list()
rmdup_stats <- list()
for (strain in strains){
  genstats[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam_rmdup/"
                                          ,strain,".genstats.txt"),sep="\t",header=F)
  
  colnames(genstats[[strain]]) <- c(	'Contig_name','Avg_fold',	'Length',	'Ref_GC',	
                                     'Covered_percent',	'Covered_bases','Plus_reads',
                                     'Minus_reads',	'Read_GC',	'Median_fold', 'Std_Dev')

    gc[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam_rmdup/"
                                      ,strain,".gcscaffold.txt"),sep="\t",header=F)
  colnames(gc[[strain]]) <- c('Contig_name',	'Length',	'A',	'C',	'G',	'T',	'N',	
                              'IUPAC','Other',	'GC')

    #genstats[[strain]]$Contig_name <- str_split(genstats[[strain]]$Contig_name,regex("_1$"),simplify = T)[,1]
    
  rmdup_stats[[strain]] <- inner_join(genstats[[strain]],gc[[strain]],by="Contig_name")


}
```

#### Read in stats for contigs scaffolded to RCC874 using RagTag

```{r}
scaffold_stats <- read.csv("data/RCC874_scaffolding_stats/merged_stats_clean_numeric.txt")
```

#### Calculate globals assembly stats including N50 and total assembled length
```{r}
calculate_stats <- function(data,prefix){
  n50s <- c()
lengths <- c()
contig_counts <- c()
mins <- c()
maxs <- c()
l50s <- c()
for (strain in strains){ 
contig_lengths <- data[[strain]]$Length.x

contig_counts <- c(contig_counts,length(contig_lengths))
total_assembled_length=sum(contig_lengths)
contig_lengths <- sort(contig_lengths,decreasing=TRUE)


sum <- 0
count <- 0
for (length in contig_lengths){
  sum <- sum+length
  count <- count+1
  if (sum>=total_assembled_length/2){
    l50s <- c(l50s,count)
    n50s <- c(n50s,as.numeric(length))
    lengths <- c(lengths,total_assembled_length)
    maxs <- c(maxs,contig_lengths[1])
    mins <- c(mins,contig_lengths[length(contig_lengths)])
    break
  }
}
if (length(contig_lengths)==0){
  l50s <- c(l50s,0)
  n50s <- c(n50s,0)
  lengths <- c(lengths,0)
  maxs <- c(maxs,0)
  mins <- c(mins,0)
}

}
stats <- data.frame(n50s,l50s,lengths,contig_counts,mins,maxs)
rownames(stats) <- strains
colnames(stats) <- c(paste0(prefix,'_N50'),paste0(prefix,'_L50'),  paste0(prefix,'_Total assembled length'),paste0(prefix,'_Contig count'),
                            paste0(prefix,'_Min contig length'),paste0(prefix,'_Max contig length'))
return(stats)
}

global_stats <- calculate_stats(dup_stats,"dup")
```

#### Add in rmdup stats
```{r}
global_stats <- cbind(global_stats,calculate_stats(rmdup_stats,"rmdup"))
```
### Add in contam stats
```{r}
global_stats <- cbind(global_stats,calculate_stats(contam_stats,"contam"))
```


#### Add in estimated genome size stats, calculated using Genomescope
```{r}
name_translation <- read.table('data/genomescope/illumina-run-conversions.txt',sep='')
temp <- name_translation$V2
names(temp) <- name_translation$V1
name_translation <- temp
```

```{r}
global_stats$genome_haploid_length <- seq(1,nrow(global_stats))
global_stats$genome_unique_length <- seq(1,nrow(global_stats))
for (folder in list.dirs('data/genomescope/')){
  if (grepl('HA',folder)){
  key <- str_split(folder,'_',simplify=TRUE)[,1]
  key <- str_split(key,'/',simplify=TRUE)[,4]
  if (sum(grepl(name_translation[key],rownames(global_stats)))==1){
  temp <- read.csv(paste0(folder,"/summary.txt_fixed.csv"))
  global_stats[name_translation[key],'genome_haploid_length'] <- 
    mean(temp[2,'min'],temp[2,'max'])
  global_stats[name_translation[key],'genome_unique_length'] <- 
    mean(temp[4,'min'],temp[4,'max'])
  }
  }
}
#print(temp)
print(global_stats)
```
We have 15 strain genomes in total, 14 of which have a predicted unique and haploid 
genome length using genomescope.  This algorithm did not converge for strain RCC1215.
```{r}
global_stats$strain <- rownames(global_stats)
global_stats <- join(global_stats,scaffold_stats)
```


#### Genomescope predicted haploid genome length vs N50
```{r}

colors <- c("Dup" = "black", "Rmdup" = "blue","Contam"="red","Scaffold"="green")

ggplot(global_stats,aes(x=scaffold_N50,y=genome_haploid_length,label=strain,color="Scaffold"))+
  geom_point()+
  geom_point(aes(x=rmdup_N50,y=genome_haploid_length, color = "Rmdup"))+
  geom_point(aes(x=contam_N50,y=genome_haploid_length, color = "Contam"))+
  geom_point(aes(x=dup_N50,y=genome_haploid_length, color = "Dup"))+
  theme_bw()+
  scale_x_log10()+
  #geom_smooth(method=lm,colour="black")+
  ylim(5e7,1.5e8)+
  ylab("Haploid genome length")+
  xlab("Contig N50")+
  labs(caption=str_wrap("Figure 1.  Relationship between predicted haploid genome 
                        length and contig N50 for 14 Emiliania huxleyi  
                        genomes.",75))+
    geom_text_repel(size=3,color="black")+
  geom_text_repel(
                aes(x = rmdup_N50,
                    y = genome_haploid_length),
                color= "black",size=3)+
  theme(plot.caption = element_text(hjust = 0,size=12))+
  labs(color = "Legend")+   scale_color_manual(values = colors)


ggsave("plots/haploid_genome_length_vs_N50.png")
```


#### Genomescope predicted unique genome length vs N50
```{r}


#global_stats$Strain <- rownames(global_stats)

ggplot(global_stats,aes(x=scaffold_N50,y=genome_unique_length,label=strain,color="Scaffold"))+
  geom_point()+
  geom_point(aes(x=rmdup_N50,y=genome_unique_length, color = "Rmdup"))+
  geom_point(aes(x=contam_N50,y=genome_unique_length, color = "Contam"))+
  geom_point(aes(x=dup_N50,y=genome_unique_length, color = "Dup"))+
  theme_bw()+
  scale_x_log10()+
  #geom_smooth(method=lm,colour="black")+
  ylim(5e7,1e8)+
  ylab("Unique genome length")+
  xlab("Contig N50")+
  labs(caption=str_wrap("Figure 2.  Relationship between predicted unique genome 
                        length and contig N50 for 14 Emiliania huxleyi  
                        genomes.",75))+
    geom_text_repel(size=3,color="black")+
  geom_text_repel(
                aes(x = rmdup_N50,
                    y = genome_unique_length),
                color= "black",size=3)+
  theme(plot.caption = element_text(hjust = 0,size=12))+
  labs(color = "Legend")+   scale_color_manual(values = colors)


ggsave("plots/unique_genome_length_vs_N50.png")


```


#### Contig length distribution for each strain
```{r}


stats_joined <- ldply(contam_stats,rbind)
stats_joined$.id <- factor(stats_joined$.id,levels=c("CCMP371",  "CCMP375",  "CCMP377",  "CCMP1280", "RCC874",   "RCC914",   "RCC1222",  "RCC1239",  "RCC1256",  "RCC3492",  "RCC3963",  "RCC6071",  "RCC6856", "RCC1212","RCC1215"))
contam <- ggplot(stats_joined,aes(x=Length.x))+geom_histogram()+scale_x_log10()+
           theme_bw()+
  xlab("Contig size")+
  ylab("Count")+facet_wrap(vars(.id))+
  ylim(0,2500)+
  theme(plot.caption = element_text(hjust = 0,size=12),
        axis.text.x = element_text(angle = 90))


stats_joined <- ldply(dup_stats,rbind)
stats_joined$.id <- factor(stats_joined$.id,levels=c("CCMP371",  "CCMP375",  "CCMP377",  "CCMP1280", "RCC874",   "RCC914",   "RCC1222",  "RCC1239",  "RCC1256",  "RCC3492",  "RCC3963",  "RCC6071",  "RCC6856", "RCC1212","RCC1215"))
dup <- ggplot(stats_joined,aes(x=Length.x))+geom_histogram()+scale_x_log10()+
           theme_bw()+
  xlab("Contig size")+
  ylab("Count")+facet_wrap(vars(.id))+
  ylim(0,2500)+
  theme(plot.caption = element_text(hjust = 0,size=12), axis.text.x = element_text(angle = 90))

stats_joined <- ldply(rmdup_stats,rbind)
stats_joined$.id <- factor(stats_joined$.id,levels=c("CCMP371",  "CCMP375",  "CCMP377",  "CCMP1280", "RCC874",   "RCC914",   "RCC1222",  "RCC1239",  "RCC1256",  "RCC3492",  "RCC3963",  "RCC6071",  "RCC6856", "RCC1212","RCC1215"))
rmdup <- ggplot(stats_joined,aes(x=Length.x))+geom_histogram()+scale_x_log10()+
           theme_bw()+
  xlab("Contig size")+
  ylab("Count")+facet_wrap(vars(.id))+
  ylim(0,2500)+
  theme(plot.caption = element_text(hjust = 0,size=12),
        axis.text.x = element_text(angle = 90))

combined <- plot_grid(contam,dup,rmdup, labels = c('Contam','Dup', 'Rmdup'), label_size = 12,ncol=3,vjust=75)
title <- ggdraw() + draw_label("Figure 3. Contig length distribution for 15 Emiliania huxleyi genomes",size=10)+theme(panel.background = element_rect(fill = 'white',color="white"))
plot_grid(combined,title, ncol=1, rel_heights=c(1, 0.1))
ggsave("plots/contig_length_distributions.png",width=20,height=10)

```




#### Scatterplots of fold coverage vs GC percentage (each point represents a contig)
```{r }
stats_joined <- ldply(contam_stats,rbind)
stats_joined$.id <- factor(stats_joined$.id,levels=c("CCMP371",  "CCMP375",  "CCMP377",  "CCMP1280", "RCC874",   "RCC914",   "RCC1222",  "RCC1239",  "RCC1256",  "RCC3492",  "RCC3963",  "RCC6071",  "RCC6856", "RCC1212","RCC1215"))
contam <- ggplot(data = stats_joined, aes(x = GC,y=Avg_fold)) +
  geom_point(size=0.5)+
  ylim(0,300)+
  xlim(0.25,0.9)+
  #scale_y_log10()+
  facet_wrap(vars(.id))+
             theme_bw()+
  xlab("Contig GC%")+
  ylab("Average Fold Coverage")+
  theme(plot.caption = element_text(hjust = 0,size=12), axis.text.x = element_text(angle = 90))
nrow(stats_joined)

stats_joined <- ldply(dup_stats,rbind)
stats_joined$.id <- factor(stats_joined$.id,levels=c("CCMP371",  "CCMP375",  "CCMP377",  "CCMP1280", "RCC874",   "RCC914",   "RCC1222",  "RCC1239",  "RCC1256",  "RCC3492",  "RCC3963",  "RCC6071",  "RCC6856", "RCC1212","RCC1215"))
dup <- ggplot(data = stats_joined, aes(x = GC,y=Avg_fold)) +
  geom_point(size=0.5)+
  ylim(0,300)+
  xlim(0.25,0.9)+
  #scale_y_log10()+
  facet_wrap(vars(.id))+
             theme_bw()+
  xlab("Contig GC%")+
  ylab("Average Fold Coverage")+
  theme(plot.caption = element_text(hjust = 0,size=12), axis.text.x = element_text(angle = 90))
nrow(stats_joined)

stats_joined <- ldply(rmdup_stats,rbind)
stats_joined$.id <- factor(stats_joined$.id,levels=c("CCMP371",  "CCMP375",  "CCMP377",  "CCMP1280", "RCC874",   "RCC914",   "RCC1222",  "RCC1239",  "RCC1256",  "RCC3492",  "RCC3963",  "RCC6071",  "RCC6856", "RCC1212","RCC1215"))
rmdup <- ggplot(data = stats_joined, aes(x = GC,y=Avg_fold)) +
  geom_point(size=0.5)+
  ylim(0,300)+
  xlim(0.25,0.9)+
  #scale_y_log10()+
  facet_wrap(vars(.id))+
             theme_bw()+
  xlab("Contig GC%")+
  ylab("Average Fold Coverage")+
  theme(plot.caption = element_text(hjust = 0,size=12),
        axis.text.x = element_text(angle = 90))
nrow(stats_joined)

combined <- plot_grid(contam,dup,rmdup, labels = c('Contam','Dup', 'Rmdup'), label_size = 12,ncol=3,vjust=75)
title <- ggdraw() + draw_label("Figure 4. Contig fold coverage vs GC percentage for 15 Emiliania huxleyi genomes",size=10)+theme(panel.background = element_rect(fill = 'white',color="white"))
plot_grid(combined,title, ncol=1, rel_heights=c(1, 0.1))

ggsave("plots/coverage_vs_GC.png",height=10,width=20)
```


#### Genomescope predicted unique genome length vs assembled length
```{r}
global_stats$rmdup_total_assembled_length <- global_stats[,"rmdup_Total assembled length"]
global_stats$dup_total_assembled_length <- global_stats[,"dup_Total assembled length"]
global_stats$contam_total_assembled_length <- global_stats[,"contam_Total assembled length"]

ggplot(global_stats,aes(x=rmdup_total_assembled_length,y=genome_unique_length,label=strain,color="Rmdup"))+
  geom_point()+
  geom_point(aes(x=dup_total_assembled_length,y=genome_unique_length, color = "Dup"))+
  geom_point(aes(x=contam_total_assembled_length,y=genome_unique_length, color = "Contam"))+
  geom_point(aes(x=assembled_scaffold_length,y=genome_unique_length, color = "Scaffold"))+
  theme_bw()+
  geom_abline(color="black",slope=1,linetype = "dashed")+
    geom_abline(color="black",slope=0.5,linetype = "dashed")+
  xlim(0,4e+8)+
  ylim(0,1.5e+8)+
  ylab("Unique genome length")+
  xlab("Total assembled length")+
  labs(caption=str_wrap("Figure 5.  Relationship between predicted unique genome
                        length and total assembled length for 14 Emiliania huxleyi 
                        genomes.",75))+
    geom_text_repel(size=3,nudge_x=10,min.segment.length=0,color="black")+
    annotate("text",x=1e8,y=1e8,label="y=x",color="red",size=5)+
      annotate("text",x=2.3e8,y=1e8,label="y=0.5x",color="red",size=5)+
  theme(plot.caption = element_text(hjust = 0,size=12))+
  labs(color = "Legend")+
  scale_color_manual(values = colors)



ggsave("plots/unique_genome_length_vs_total_assembled_length.png")
```

```{r}



ggplot(global_stats,aes(x=rmdup_total_assembled_length,y=genome_haploid_length,label=strain,color="Rmdup"))+
  geom_point()+
  geom_point(aes(x=dup_total_assembled_length,y=genome_haploid_length, color = "Dup"))+
  geom_point(aes(x=contam_total_assembled_length,y=genome_haploid_length, color = "Contam"))+
    geom_point(aes(x=assembled_scaffold_length,y=genome_haploid_length, color = "Scaffold"))+
  theme_bw()+
  geom_abline(color="black",slope=1,linetype = "dashed")+
    geom_abline(color="black",slope=0.5,linetype = "dashed")+
  xlim(0,4e+8)+
  ylim(0,1.5e+8)+
  ylab("Haploid genome length")+
  xlab("Total assembled length")+
  labs(caption=str_wrap("Figure 6.  Relationship between predicted haploid genome
                        length and total assembled length for 14 Emiliania huxleyi 
                        genomes.",75))+
    geom_text_repel(size=3,nudge_x=10,min.segment.length=0,color="black")+
    annotate("text",x=1e8,y=1e8,label="y=x",color="red",size=5)+
      annotate("text",x=1.3e8,y=5e7,label="y=0.5x",color="red",size=5)+
  theme(plot.caption = element_text(hjust = 0,size=12))+
  labs(color = "Legend")+
  scale_color_manual(values = colors)

ggsave("plots/haploid_genome_length_vs_total_assembled_length.png")
```









