---
title: "all_strain_genome_stats"
output: pdf_document
date: "2023-07-12"
---
## Assembly stats for 13 Emiliania huxleyi genomes

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(stringr)
library(ggrepel)
library(plyr)
```


#### Read in assembly stats data

```{r read_in}
genstats <- list()
gc <- list()
stats <- list()
strains <- as.character(read.csv('data/strains.csv',header=FALSE))
for (strain in strains){
  genstats[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam/"
                                          ,strain,".genstats.txt"),sep="\t",header=F)
  colnames(genstats[[strain]]) <- c(	'Contig_name','Avg_fold',	'Length',	'Ref_GC',	
                                     'Covered_percent',	'Covered_bases','Plus_reads',
                                     'Minus_reads',	'Read_GC',	'Median_fold', 'Std_Dev')

    gc[[strain]] <- read.table(paste0("data/2023-genome-stats/stats_after_pilon_round_2_decontam/"
                                      ,strain,".gcscaffold.txt"),sep="\t",header=F)
  colnames(gc[[strain]]) <- c('Contig_name',	'Length',	'A',	'C',	'G',	'T',	'N',	
                              'IUPAC','Other',	'GC')
  stats[[strain]] <- inner_join(genstats[[strain]],gc[[strain]],by="Contig_name")
}
```

#### Calculate globals assembly stats including N50 and total assembled length
```{r}
n50s <- c()
lengths <- c()
contig_counts <- c()
mins <- c()
maxs <- c()
for (strain in strains){ 
contig_lengths <- stats[[strain]]$Length.x
contig_counts <- c(contig_counts,length(contig_lengths))
total_assembled_length=sum(contig_lengths)
contig_lengths <- sort(contig_lengths,decreasing=TRUE)

sum <- 0
for (length in contig_lengths){
  sum <- sum+length
  if (sum>=total_assembled_length/2){
    n50s <- c(n50s,as.numeric(length))
    lengths <- c(lengths,total_assembled_length)
    maxs <- c(maxs,contig_lengths[1])
    mins <- c(mins,contig_lengths[length(contig_lengths)])
    break
  }
}

}
global_stats <- data.frame(n50s,lengths,contig_counts,mins,maxs)
rownames(global_stats) <- strains
colnames(global_stats) <- c('N50','Total assembled length','Contig count',
                            'Min contig length','Max contig length')
```
#### Add in estimated genome size stats, calculated using Genomescope
```{r}
name_translation <- read.table('data/genomescope/illumina-run-conversions.txt',sep='')
temp <- name_translation$V2
names(temp) <- name_translation$V1
name_translation <- temp
```
```{r}
global_stats$genome_haploid_length <- seq(1,nrow(global_stats))
global_stats$genome_unique_length <- seq(1,nrow(global_stats))
for (folder in list.dirs('data/genomescope/')){
  if (grepl('HA',folder)){
  key <- str_split(folder,'_',simplify=TRUE)[,1]
  key <- str_split(key,'/',simplify=TRUE)[,4]
  if (sum(grepl(name_translation[key],rownames(global_stats)))==1){
  temp <- read.csv(paste0(folder,"/summary.txt_fixed.csv"))
  global_stats[name_translation[key],'genome_haploid_length'] <- 
    mean(temp[2,'min'],temp[2,'max'])
  global_stats[name_translation[key],'genome_unique_length'] <- 
    mean(temp[4,'min'],temp[4,'max'])
  }
  }
}
#print(temp)
print(global_stats)
```

#### Genomescope predicted haploid genome length vs N50
```{r}
global_stats$Strain <- rownames(global_stats)
ggplot(global_stats,aes(x=N50,y=genome_haploid_length,label=Strain))+
  geom_point()+
  theme_bw()+
  scale_x_log10()+
  geom_smooth(method=lm,colour="black")+
  ylab("Haploid genome length")+
  xlab("Contig N50")+
  labs(caption=str_wrap("Figure 1.  Relationship between predicted haploid genome 
                        length and contig N50 for 13 Emiliania huxleyi  
                        genomes.",75))+
    geom_text_repel(size=3)+
  theme(plot.caption = element_text(hjust = 0,size=12))
ggsave("plots/haploid_genome_length_vs_N50.png")
```
#### Genomescope predicted unique genome length vs N50
```{r}

global_stats$Strain <- rownames(global_stats)
ggplot(global_stats,aes(x=N50,y=genome_unique_length,label=Strain))+
  geom_point()+
  theme_bw()+
  scale_x_log10()+
  geom_smooth(method=lm,colour="black")+
  ylab("Unique genome length")+
  xlab("Contig N50")+
  labs(caption=str_wrap("Figure 2.  Relationship between predicted unique genome
                        length and contig N50 for 13 Emiliania huxleyi 
                        genomes.",75))+
    geom_text_repel(size=3)+
  theme(plot.caption = element_text(hjust = 0,size=12))
ggsave("plots/unique_genome_length_vs_N50.png")
```



#### Contig length distribution for each strain
```{r}
stats_joined <- ldply(stats,rbind)
ggplot(stats_joined,aes(x=Length.x))+geom_histogram()+scale_x_log10()+
           theme_bw()+
  xlab("Contig size")+
  ylab("Count")+facet_wrap(vars(.id))+
  labs(caption=str_wrap("Figure 3. Contig length distribution for 13 Emiliania 
                        huxleyi genomes.",75))+
  theme(plot.caption = element_text(hjust = 0,size=12))

ggsave("plots/contig_length_distributions.png")
```



```{r,include=FALSE}
  
stats_joined["Normalized_fold_coverage"] <- stats_joined["Avg_fold"]/mean(stats_joined$Avg_fold)

```

#### Scatterplots of fold coverage vs GC percentage (each point represents a contig)
```{r }

ggplot(data = stats_joined, aes(x = GC,y=Avg_fold)) +
  geom_point(size=0.5)+ylim(0,300)+facet_wrap(vars(.id))+
             theme_bw()+
  xlab("Contig GC%")+
  ylab("Average Fold Coverage")+
  labs(caption=str_wrap("Figure 4. Contig Fold coverage vs GC percentage for 13 
                        Emiliana huxleyi genomes.",75))+
  theme(plot.caption = element_text(hjust = 0,size=12))
ggsave("plots/coverage_vs_GC.png")
```


#### Genomescope predicted unique genome length vs assembled length
```{r}

global_stats$total_assembled_length <- global_stats[,"Total assembled length"]
ggplot(global_stats,aes(x=total_assembled_length,y=genome_unique_length,label=Strain))+
  geom_point()+
  theme_bw()+
  geom_smooth(method=lm,colour="black")+
  ylab("Unique genome length")+
  xlab("Total assembled length")+
  labs(caption=str_wrap("Figure 5.  Relationship between predicted unique genome
                        length and total assembled length for 13 Emiliania huxleyi 
                        genomes.",75))+
    geom_text_repel(size=3)+
  theme(plot.caption = element_text(hjust = 0,size=12))
ggsave("plots/unique_genome_length_vs_total_assembled_length.png")
```

```{r}


ggplot(global_stats,aes(x=total_assembled_length,y=genome_haploid_length,label=Strain))+
  geom_point()+
  theme_bw()+
  geom_smooth(method=lm,colour="black")+
  ylab("Haploid genome length")+
  xlab("Total assembled length")+
  labs(caption=str_wrap("Figure 6.  Relationship between predicted haploid genome
                        length and total assembled length for 13 Emiliania huxleyi 
                        genomes.",75))+
    geom_text_repel(size=3)+
  theme(plot.caption = element_text(hjust = 0,size=12))
ggsave("plots/haploid_genome_length_vs_total_assembled_length.png")
```

```{r include=FALSE}

temp <- stats_joined
#temp <- filter(temp,Avg_fold<50)
temp <- mutate(temp, bin=cut_width(GC, width=0.01, boundary=0) )

ggplot(temp, aes(x=bin, y=Avg_fold)) +geom_boxplot()+
  labs(caption=str_wrap(paste0("Fold coverage vs GC percentage for strain ",strain),75))+
  facet_wrap(vars(.id))+ylim(0,200)+
  theme(plot.caption = element_text(hjust = 0,size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```






